---
title: "DATA605 ASSIGNMENT 10"
author: "Fan Xu"
date: "11/1/2020"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load library, message=FALSE, warning=FALSE}
library(tidyverse)


```

# Question 1

**Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars. A guard agrees to make a series of bets with him. If Smith bets A dollars, he wins A dollars with probability .4 and loses A dollars with probability .6.**

**Find the probability that he wins 8 dollars before losing all of his money if**

## Part (a)
**he bets 1 dollar each time (timid strategy).**

**Anwer：**

$Let\;P(win)=q=0.4\;and\;P(lose)=q=1-p=0.6$

$P(win\;8\;dollars|bets\;1\;dollar\;each\;time)=$


according to the formula of 'Gambler's Ruin':

$P_{i}(n)=\frac{1-(\frac{q}{p})^t}{1-(\frac{q}{p})^n}\;if\;p\neq q$

$\because$ 1 dollar each time and wins 8 dollars

$\therefore\;i=1\;and\;n=8$

$\therefore\;P_{1}(8)=\frac{1-(\frac{0.6}{0.4})^i}{1-(\frac{0.6}{0.4})^n}=0.02030135$

```{r}
p = 0.4
q = 1-p

(1-(q/p)^1)/(1-(q/p)^8)
```

## Part (b)
**he bets, each time, as much as possible but not more than necessary to bring his fortune up to 8 dollars (bold strategy).**

**Answer:**

Using bold strategy, he fails to get out on bail whenever he loss any bets, therefore there is only one scenario where he can win 8 dollars:

a. Win the first time, get 2 dollars,

b. Win the second time, get 4 dollars,

c. Win the third time, get $8 dollars.

The probability of wining 3 times consectively is $P(3 wins consectively)=0.4^3=0.064$

```{r}
0.4^3

```





## Part (c)
**Which strategy gives Smith the better chance of getting out of jail?**

**Answer:**

becuase 0.064 > 0.0203, therefore using bold strategy giving Smith better chance of getting out of jail.
