---
title: "DATA605 ASSIGNMENT 8"
author: "Fan Xu"
date: "10/18/2020"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load library, message=FALSE, warning=FALSE}
library(tidyverse)


```

# Question 11 (Pg 303)

**A company buys 100 lightbulbs, each of which has an exponential lifetime of 1000 hours. What is the expected time for the first of these bulbs to burn out? (See Exercise 10.)**


**Answer:**

$Let \;\lambda=\frac{1}{1000}\;,n=100\;and\;X_{min}=m$

$\because P(X_{min}=m)=1-P(X_{i}>=m\;for\;i\in [1,100])$ <br>
$=1-(e^{-\lambda m})^n)$ <br>
$=1-e^{-\frac{100}{1000}m}$ <br>
$=1-e^{-\frac{1}{10}m}$

$\therefore E[m] = \frac{1}{\frac{1}{10}}=10$



# Question 1 (Pg 320-321)

**Let X be a continuous random variable with mean μ = 10 and variance $\sigma^{2} = \frac{100}{3}$. Using Chebyshev’s Inequality, find an upper bound for the following probabilities.**

**Answer**
Let $\mu = 10\;and\;\sigma=\sqrt\frac{100}{3}$,

and according to Chebyshev's Inqeuqlity, $P(\left | X-\mu \right |\geq k\sigma)\leq\frac{1}{K^{2}}$


## (a) 
**$P(\left | X-10 \right |\geq2).$**

$\because k=\frac{2}{\sigma}\;and\;k^{2}=\frac{4}{\sigma^{2}}=\frac{4}{\frac{100}{3}}=\frac{3}{25}$

$\therefore P(\left | X-10 \right |\geq2)\leq min(\frac{1}{k^{2}},1)= min(\frac{25}{3},1)=1$

$\therefore P(\left | X-10 \right |\geq2)\leq 1$


## (b) 
**$P(\left | X-10 \right |\geq5).$**

$\because k=\frac{5}{\sigma}\;and\;k^{2}=\frac{25}{\sigma^{2}}=\frac{25}{\frac{100}{3}}=\frac{3}{4}$

$\therefore P(\left | X-10 \right |\geq5)\leq min(\frac{1}{k^{2}},1)= min(\frac{4}{3},1)=1$

$\therefore P(\left | X-10 \right |\geq5)\leq 1$

## (c) 
**$P(\left | X-10 \right |\geq9).$**

$\because k=\frac{9}{\sigma}\;and\;k^{2}=\frac{81}{\sigma^{2}}=\frac{81}{\frac{100}{3}}=\frac{243}{100}$

$\therefore P(\left | X-10 \right |\geq9)\leq min(\frac{1}{k^{2}},1)= min(\frac{100}{243},1)=\frac{100}{243}$

$\therefore P(\left | X-10 \right |\geq9)\leq \frac{100}{243}\approx0.411523$

## (d) 
**$P(\left | X-10 \right |\geq20).$**

$\because k=\frac{20}{\sigma}\;and\;k^{2}=\frac{400}{\sigma^{2}}=\frac{400}{\frac{100}{3}}=12$

$\therefore P(\left | X-10 \right |\geq20)\leq min(\frac{1}{k^{2}},1)= min(\frac{1}{12},1)=\frac{1}{212}$

$\therefore P(\left | X-10 \right |\geq20)\leq \frac{1}{12}\approx0.083333$


# Question 14 (Pg 303)

***14 Assume that X1 and X2 are independent random variables, each having an exponential density with parameter $\lambda$. Show that Z = X1 − X2 has density**

$$f_{Z}(z)=(\frac{1}{2}\lambda e^{-\lambda\left | z \right |})$$

**Answer**

Let $$f_{X_{i}}(x)=\left\{\begin{matrix}\lambda e^{-\lambda x}\;\;\;\;\;\;x>0\\ 0\;\;\;\;\;\;\;otherwise\end{matrix}\right.\;\;\;\;for\;i\in[1,2]$$

then let $x_{1}=x$ and $x_{1}=y$
for $z = x_{1}-x_{2} = x-y$, then $y=x-z$

$f_{Z}(z)=\int_{-\infty}^{\infty}f_{Y}(x-z)f_{X}(x)dx$

for $z\geq0$ <br>
$\because x-z>0$ <br>
$\therefore x>z$

$f_{Z}(z)=\int_{z}^{\infty}\lambda e^{-\lambda (x-z)}\lambda e^{-\lambda x}dx$ <br>
$=\int_{z}^{\infty}\lambda^{2}e^{-\lambda(2x-z)}dx$ <br>
$=\lambda^{2}e^{\lambda z}\int_{z}^{\infty}e^{-2\lambda x}dx$ <br>
$=\lambda^{2}e^{\lambda z}\frac{1}{-2\lambda}(e^{-2\lambda\infty}-e^{-2\lambda z})$ <br>
$=-\frac{1}{2}\lambda e^{\lambda z}(-e^{-2\lambda z})$ <br>
$$=\frac{1}{2}\lambda e^{-\lambda z}$$

for $z<0$ <br>
$\because x-z>0$ <br>
$\therefore x>0>z$ <br>
$\therefore x>0$

$f_{Z}(z)=\int_{0}^{\infty}\lambda e^{-\lambda (x-z)}\lambda e^{-\lambda x}dx=$ <br>
$=\int_{0}^{\infty}\lambda^{2}e^{-\lambda(2x-z)}dx$ <br>
$=\lambda^{2}e^{\lambda z}\int_{0}^{\infty}e^{-2\lambda x}dx$ <br>
$=\lambda^{2}e^{\lambda z}\frac{1}{-2\lambda}(e^{-2\lambda\infty}-e^{-2\lambda 0})$ <br>
$=\lambda^{2}e^{\lambda z}\frac{1}{-2\lambda}(0-1)$ <br>
$$=\frac{1}{2}\lambda e^{\lambda z}$$

\therefore

$$f_{Z}(z)=\left\{\begin{matrix}\frac{1}{2}\lambda e^{-\lambda z}\;\;\;\;\;\;z\geq0\\ \frac{1}{2}\lambda e^{\lambda z}\;\;\;\;\;\;\;z<0\end{matrix}\right.$$

In short:

$$f_{Z}(z)=\frac{1}{2}\lambda e^{-\lambda\left | z \right |}$$