---
title: "DATA605 ASSIGNMENT 7"
author: "Fan Xu"
date: "10/11/2020"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load library, message=FALSE, warning=FALSE}
library(tidyverse)


```

# Question 1
Let X1, X2, . . . , Xn be n mutually independent random variables, each of
which is uniformly distributed on the integers from 1 to k. Let Y denote the
minimum of the Xi’s. Find the distribution of Y .


**Answer:**

$\because x_{i}$ is uniformly distrubted on the intergers from 1 to k

$\therefore$ the pdf of $x_{i} $ is $P(X=x_{i}) = \frac{1}{k}$

$\because x_{i}'s$ are mutually independtent from each other

$\therefore$ the joint pdf of $P(All(X_{i}) > x) = P(x_{1} > x,x_{2} > x,...,x_{n} > x \;for\;x\;in\;range\;(1,k)) = (1-\frac{x}{k})^n$

$\therefore P(Min(x_{i})=1) = P(Y = 1) = 1-P(All(X_{i}) > 1)=1-(1-\frac{1}{k})^n=(1-\frac{0}{k})^n-(1-\frac{1}{k})^n$

$\therefore P(Y = 2) = 1-P(All(X_{i}) > 2) - P(Y = 1)=(1-\frac{1}{k})^n-(1-\frac{2}{k})^n$

$\therefore P(Y = 3) = 1-P(All(X_{i}) > 3) - P(Y = 1) - P(Y = 2)=(1-\frac{2}{k})^n-(1-\frac{3}{k})^n$

$...$

$\therefore P(Y = y) = 1-P(All(X_{i}) > y) - P(Y = j\;for\;j < y)=(1-\frac{y-1}{k})^n-(1-\frac{y}{k})^n$

$\therefore P(Y = y) = (\frac{k-y+1}{k})^n-(\frac{k-y}{k})^n$


# Question 2

Your organization owns a copier (future lawyers, etc.) or MRI (future doctors). This machine has a manufacturer’s expected lifetime of 10 years. This means that we expect one failure every ten years. (Include the probability statements and R Code for each part.).


## Part a
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a geometric. (Hint: the probability is equivalent to not failing during the first 8 years..)

**Answer:**

Modeling using geometric distirbution, the probability that the machine will fail after 8 years is equivalent to it will not fail in the first 8 year, therefore

$P(X = n) = (1-p)^{n-1}p$ where $p = \frac{1}{10}$

and $P(X > 8) = 1 - P(X \leq 8)$

compute in R

1. $P(x > 8)$
```{r geom P(x>8)}
p=1/10
1-pgeom(8-1,p)
```

2. $E[x] = \frac{1}{p}$
```{r geom E[x]}
1/p

```

3. $SD(x)=\sqrt{\frac{1-p}{p^2}}$
```{r geom var(x)}
sqrt((1-p)/p^2)

```

## Part b
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as an exponential.

**Answer:**

Modeling using exponential distirbution, the probability that the machine will fail after 8 years is equivalent to it will not fail in the first 8 year, therefore

$P(X <= n) = 1-e^{-\lambda x}$ where $\lambda = 1/10$

and $P(X > 8) = 1 - P(X \leq 8)$

compute in R

1. $P(x > 8)$
```{r exp P(x>8)}
lambda <- 1/10
1-pexp(8, lambda)


```

2. $E[x] = \frac{1}{\lambda}$
```{r exp E[x]}
1/lambda
```


3. $SD(x) = \sqrt{\frac{1}{\lambda^2}}$
```{r exp var(x)}
sqrt(1/lambda^2)
```

## Part c

What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a binomial. (Hint: 0 success in 8 years)


Modeling using binomial distirbution, the probability that the machine will fail after 8 years is equivalent to it fail 0 times in the first 8 year, therefore

$P(X = n) = \begin{pmatrix}n\\ x\end{pmatrix}p^x(1-p)^{n-x}$ where $p = \frac{1}{10}\;and\;n=8$

and $P(x) = \begin{pmatrix}8\\ 0\end{pmatrix}(\frac{1}{10})^0(\frac{9}{10})^{8}$

compute in R

1.$P(x=0)\;given\;p=\frac{1}{10}\;and\;n=8$
```{r binom}
p <- 1/10
n <- 8
pbinom(0,n,p)


```

2. $E[x] = np$
```{r Binom E[x]}
n*p


```

3. $SD[x]=\sqrt{np(1-p)}$
```{r Binom var(x)}
sqrt(n*p*(1-p))

```


## Part d
What is the probability that the machine will fail after 8 years?. Provide also the expected value and standard deviation. Model as a Poisson.


Modeling using Poisson distirbution, the probability that the machine will fail after 8 years is equivalent to it will fail zero times in the first 8 year, therefore

for 1 year, $P(X = n) = \frac{\lambda^{x}e^{-\lambda}}{x!}$ where $\lambda = 1/10$, and 

for 8 year, $P(X = n) = \frac{\lambda^{x}e^{-\lambda}}{x!}$ where $\lambda = 8/10$ 


compute in R

1.$P(x=0)\;given\;\lambda=\frac{8}{10}$
```{r pois}
lambda = 8/10
ppois(0,lambda)


```

2.$E[x] = \lambda$ 
```{r pois E[x]}
lambda
```


3.$SD[x] = \sqrt{\lambda}$ 
```{r pois var(x)}
sqrt(lambda)
```