---
title: "DATA605 ASSIGNMENT 14"
author: "Fan Xu"
date: "12/06/2020"
output:
  rmdformats::readthedown:
    code_folding: hide
    df_print: paged
    highlight: tango
    number_sections: yes
    smooth_scroll: yes
    theme: united
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_collapsed: yes
    toc_float: yes
  pdf_document:
    extra_dependencies:
    - geometry
    - multicol
    - multirow
  word_document:
    toc: yes
    toc_depth: '5'
theme: lumen
number_sections: yes
toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load library, message=FALSE, warning=FALSE}
library(tidyverse)
library(tidymodels)
library(gridExtra)
```



# Question 1
**This week, we'll work out some Taylor Series expansions of popular functions.**

- $f(x) = \frac{1}{(1-x)}$
- $f(x) = e^{x}$
- $f(x) = ln(1+x)$

**For each function, only consider its valid ranges as indicated in the notes when you are computing the Taylor Series expansion. Please submit your assignment as a R-Markdown document.**


## Part (a)
$f(x) = \frac{1}{(1-x)}$

**Answer:**

$f(x) = \frac{1}{(1-x)}$

Compute the first four derivatives corresponding to the $f(x)$ and their values at $x=0$:

$f'(x) = (-1)\cdot(1-x)^{-2}\cdot(-1) = (1-x)^{-2}$

$f'(0) = 1 = 1!$

$f''(x) = (-2)\cdot(1-x)^{-3}\cdot(-1) = 2!(1-x)^{-3}$

$f''(0) = 2!$

$f'''(x) = (2!)\cdot(-3)\cdot(1-x)^{-4}\cdot(-1) = 3!(1-x)^{-4}$

$f'''(0) = 3!$

$f''''(x) = (3!)\cdot(-4)\cdot(1-x)^{-5}\cdot(-1) = 4!(1-x)^{-5}$

$f''''(0) = 4!$

Therefore, we have the Taylor Series expansions:

$f(x) = (1-x)^{-1}$

$= \sum_{n=0}^{\infty} \frac{f^{(n)}(0)}{n!}x^{n}$

$= f(0) + f'(0)x + f''(0)\frac{x^{2}}{2!} + f'''(0)\frac{x^{3}}{3!} + \cdots$

$= 1 + 1!\cdot x + 2!\cdot \frac{x^{2}}{2!} + 3!\cdot \frac{x^{3}}{3!} +  4!\cdot \frac{x^{4}}{4!} + \cdots$

$= 1 + x + x^{2} + x^{3} + x^{4} + \cdots$

This function converges when $|x|<1$.


## Part (b)
$f(x) = e^{x}$

**Answer:**

$f(x) = e^{x}$

Compute the first four derivatives corresponding to the $f(x)$ and their values at $x=0$:

$f'(x) = e^{x}$

$f'(0) = 1$

$f''(x) = e^{x}$

$f''(0) = 1$

$f'''(x) = e^{x}$

$f'''(0) = 1$

$f''''(x) = e^{x}$

$f''''(0) = 1$

Therefore, we have the Taylor Series expansions:

$f(x) = e^{x}$

$= \sum_{n=0}^{\infty} \frac{f^{(n)}(0)}{n!}x^{n}$

$= f(0) + f'(0)x + f''(0)\frac{x^{2}}{2!} + f'''(0)\frac{x^{3}}{3!} + \cdots$

$= 1 + x + \frac{x^{2}}{2!} + \frac{x^{3}}{3!} + \frac{x^{4}}{4!} + \cdots$

$= \sum_{n=0}^{\infty} \frac{x^{n}}{n!}$

This function converges for all $x \in R$.


## Part (c)
$f(x) = ln(1+x)$

**Answer:**

$f(x) = ln(1+x)$

Compute the first four derivatives corresponding to the $f(x)$ and their values at $x=0$:

$f'(x) = (1+x)^{-1}$

$f'(0) = 1 = (-1)^{(1+1)} \cdot 0!$

$f''(x) = (-1) \cdot (1+x)^{-2}$

$f''(0) = -1 = (-1)^{(2+1)} \cdot 1!$

$f'''(x) = (-1) \cdot (-2) \cdot (1+x)^{-3}$

$f'''(0) = 2  = (-1)^{(3+1)} \cdot 2!$

$f''''(x) = (-1) \cdot (-2) \cdot (-3) (1+x)^{-4}$

$f''''(0) = -6 = (-1)^{(4+1)} \cdot 3!$

Therefore, we have the Taylor Series expansions:

$f(x) = ln(1+x)$

$= \sum_{n=0}^{\infty} \frac{f^{(n)}(0)}{n!}x^{n}$

$= f(0) + f'(0)x + f''(0)\frac{x^{2}}{2!} + f'''(0)\frac{x^{3}}{3!} + \cdots$

$= 0 + 1 + (-1)\frac{x^{2}}{2!} + (2)\frac{x^{3}}{3!} + (-6)\frac{x^{4}}{4!} + \cdots$

$= \sum_{n=1}^{\infty} (-1)^{n+1}(n-1)!\frac{x^{n}}{n!}$

$= \sum_{n=1}^{\infty} \frac{(-1)^{n+1}\cdot x^{n}}{n}$

This function converges for $|x|<1$.
